---
title: "Validation_data_filtering"
output: html_document
---

* Subset validation data to only the ones used
  1. Drug target references: 
    + Exclude overlapping DrugBank/Target Central/other dataset entries
    + Include only Kinase targets
  2. Filter CREEDS drug perturbation expression data:
    + Include only drugs from filtered drug-target list
* Current limitations:
    + Need a better identifier than Smiles to match up and filter drugs 


# Load packages & Subset GMT functions
```{r}
setwd("~/Desktop/x2k/Validation/")
library(cogena)

subset_GMT <- function(GMTfile, subset_names){
  count=1
  CREEDS_subset <- list()
  booleans = logical(0)
  for(i in GMTfile){
    if( names(GMTfile[count]) %in% subset_names ){
      booleans <- c(booleans, TRUE)
      #print("Add to GMT subset")
    } else{
      booleans <-c(booleans, FALSE)
      #print("Experiment not in drug-target list.")
      }
    count = count+1
  }
  GMT_subset <-GMTfile[booleans]
  return(GMT_subset)
}


# Include only kinases using Kinome database *******
library(readxl)
kinome <- read_excel("~/Desktop/x2k/General_Resources/Kinase.com/Kinome_Hsap.xls")
   
 GMT_kinome_filter <- function(GMTfile, kinome){
   booleans <-logical(0)
   count = 1
   for(i in GMTfile){
     if( sum(GMTfile[count][[1]] %in% kinome$Name) > 0 ){
      booleans <- c(booleans, TRUE)
      #print("Drug added to GMT kinome subset")
     } else{
      booleans <-c(booleans, FALSE)
      #print("Drug does not have any kinases as targets")
     }
    count = count+1
   }
  GMT_kinome <- GMTfile[booleans]
  return(GMT_kinome)
 }

```




# [Dataset 1] CREEDS_manual_single_drug_perturbations / ...
```{r}
#********** Import CREEDS drug perturbation transcriptome data: **********#
  CREEDS.ManualDrugs <-gmt2list("Perturbation_Data/CREEDS/single_drug_perturbations-v1.0.gmt")
  # Import CREEDS_drugs metadata
  CREEDS.ManualDrugs_meta <-read.csv("Perturbation_Data/CREEDS/CREEDS_metadata_single_drug_perturbations-v1.0.csv")
```

## [1.A]... / DrugBank.TargetCentral + kinome
```{r}
#********** Import combined Drugbank + Target Central drug target list **********#
DrugBank.TargetCentral <- DrugBank.TargetCentral_orig <- read.csv("Drug-Target_Data/11-17-17_WithDrugBankIDs/3_Union_EdgeList_v1_11-17-17.csv")


# Include only kinases using Kinome database *******
  DB.TC_kinome_filter <- DrugBank.TargetCentral$TargetID_Entrez %in% kinome$Entrez_GeneID
  DrugBank.TargetCentral <- DrugBank.TargetCentral[DB.TC_kinome_filter,]
    

# [A] DrugBank-TargetCentral filter
  # Filter the CREEDS metadata to just the drugs we have targets for
    #DB.TC_filter <-CREEDS.ManualDrugs_meta$smiles %in% DrugBank.TargetCentral$DrugSMILES
    DB.TC_filter <-CREEDS.ManualDrugs_meta$drugbank_id %in% DrugBank.TargetCentral$DrugID_DrugBank

  CREEDS_meta.filter <- CREEDS.ManualDrugs_meta[DB.TC_filter,]
  remodified_names <- c(paste(CREEDS_meta.filter$drug_name,"-dn",sep=""),
                        paste(CREEDS_meta.filter$drug_name,"-up",sep="") )
  
# Filter CREEDS data
CREEDS_manualDrugs_DB.TC <-subset_GMT(CREEDS.ManualDrugs, remodified_names)
length(CREEDS_manualDrugs_DB.TC)
# Write filtered list to GMT file
dir="Perturbation_Data/CREEDS/Filtered_GMTs/CREEDS_Manual.Drugs_DG.TC_filtered.gmt"
if(file.exists(dir)){file.remove(dir) } else{print("Doesn't exist")}
gmtlist2file(CREEDS_manualDrugs_DB.TC, dir)    
```

## [1.B] ... / Drug Repurposing Hub
```{r}
#********** Import combined Drug Repurposing Hub target list **********#
DrugRepurposingHub <-read.table("Drug-Target_Data/Drug Repurposing Hub/Repurposing_Hub_export.txt", 
                                  header=T, sep="\t", fill = T)
  # Get rid of any drugs that don't have targets
  DrugRepurposingHub <- subset(DrugRepurposingHub, Target!="") 
  write.table(DrugRepurposingHub, "Drug-Target_Data/Drug Repurposing Hub/DRH_validation.txt", row.names=FALSE)
  
  
# [B] Drug Repurposing Hub filter
  DRH_filter <-CREEDS.ManualDrugs_meta$smiles %in% DrugRepurposingHub$SMILES

  CREEDS_meta.filter <- CREEDS.ManualDrugs_meta[DRH_filter,]
  remodified_names <- c(paste(CREEDS_meta.filter$drug_name,"-dn",sep=""),
                        paste(CREEDS_meta.filter$drug_name,"-up",sep="") )
  
# Filter CREEDS data
CREEDS_manualDrugs_DRH <-subset_GMT(CREEDS.ManualDrugs, remodified_names)
length(CREEDS_manualDrugs_DRH)

# Write filtered list to GMT file
dir = "Perturbation_Data/CREEDS/Filtered_GMTs/CREEDS_Manual.Drugs_DRH_filtered.gmt"
if(file.exists(dir)){file.remove(dir) } else{print("Doesn't exist")}
gmtlist2file(CREEDS_manualDrugs_DRH, dir)  
```



# [Dataset 2] CREEDS_DrugMatrix / ...
```{r}
#********** Import CREEDS DrugMatrix perturbation transcriptome data: **********#
  CREEDS_DrugMatrix <- gmt2list("Perturbation_Data/CREEDS/single_drug_perturbations-DrugMatrix.gmt")
  # Import CREEDS_DrugMatrix metadata
  CREEDS_DrugMatrix_meta <- read.csv("Perturbation_Data/CREEDS/single_drug_perturbations-DrugMatrix.csv")
```

## [2.A]... / DrugBank.TargetCentral + kinome
```{r}
#********** Import combined Drugbank + Target Central drug target list **********#
DrugBank.TargetCentral <- DrugBank.TargetCentral_orig <- read.csv("Drug-Target_Data/11-17-17_WithDrugBankIDs/3_Union_EdgeList_v1_11-17-17.csv")
  # Include only kinases using Kinome database *******
    kinome_filter <- DrugBank.TargetCentral$TargetID_Entrez %in% kinome$Entrez_GeneID
    DrugBank.TargetCentral <- DrugBank.TargetCentral[kinome_filter,]
    
  
# [A] DrugBank-TargetCentral filter
  # Filter the CREEDS metadata to just the drugs we have targets for
  DB.TC_filter <-CREEDS_DrugMatrix_meta$drugbank_id %in% DrugBank.TargetCentral$DrugID_DrugBank
  CREEDS_meta.filter <- CREEDS_DrugMatrix_meta[DB.TC_filter,]
  remodified_names <- c(paste(CREEDS_meta.filter$drug_name,"-dn",sep=""),
                        paste(CREEDS_meta.filter$drug_name,"-up",sep="") )
  
# Filter CREEDS data
CREEDS_DrugMatrix_DB.TC <-subset_GMT(CREEDS_DrugMatrix, remodified_names)
length(CREEDS_DrugMatrix_DB.TC)
# Write filtered list to GMT file
dir = "Perturbation_Data/CREEDS/Filtered_GMTs/CREEDS_DrugMatrix_DG.TC_filtered.gmt"
if(file.exists(dir)){file.remove(dir) } else{print("Doesn't exist")}
gmtlist2file(CREEDS_manualDrugs_DRH, dir)  
```

## [2.B] ... / Drug Repurposing Hub
```{r}
#********** Import combined Drug Repurposing Hub target list **********#
DrugRepurposingHub <-read.table("Drug-Target_Data/Drug Repurposing Hub/Repurposing_Hub_export.txt", 
                                  header=T, sep="\t", fill = T)
  # Get rid of any drugs that don't have targets
  DrugRepurposingHub <- subset(DrugRepurposingHub, Target!="") 
  
  
# [B] Drug Repurposing Hub filter
  DRH_filter <-CREEDS_DrugMatrix_meta$smiles %in% DrugRepurposingHub$SMILES
  CREEDS_meta.filter <- CREEDS_DrugMatrix_meta[DRH_filter,]
  remodified_names <- c(paste(CREEDS_meta.filter$drug_name,"-dn",sep=""),
                        paste(CREEDS_meta.filter$drug_name,"-up",sep="") )
  
# Filter CREEDS data
CREEDS_DrugMatrix_DRH <-subset_GMT(CREEDS_DrugMatrix, remodified_names)
length(CREEDS_DrugMatrix_DRH)

# Write filtered list to GMT file
dir = "Perturbation_Data/CREEDS/Filtered_GMTs/CREEDS_DrugMatrix_DRH_filtered.gmt"
if(file.exists(dir)){file.remove(dir) } else{print("Doesn't exist")}
gmtlist2file(CREEDS_manualDrugs_DRH, dir)  
```



## [Dataset 3] DrugBank_Moshe / kinome
* The DrugBank targets (under the Mechanisms of Action section) are collected from many different sources (Wikipedia, PubChem, KEGG, etc.), not exclusively drug perturbation data. Consequently, DrugBank data could be very noisy compared to a controlled drug perturbation dataset.

```{r}
#********** Import combined Drugbank drug target list **********#
DrugBank <- gmt2list("Perturbation_Data/DrugBank/drugbank_gene_set_2017_11.gmt")
  
# Link to DrugBank's metadata provided on their website
    #drugbank_vocab <-read.csv("Drug-Target_Data/drugbank vocabulary.csv")
  # Add info in drugbank_vocab to Moshe's DrugBank data
  #drugbank_vocab.filt <-drugbank_vocab[drugbank_vocab$Common.name %in% names(DrugBank), ]
  #DBoverlap_names <- drugbank_vocab.filt$Common.name
# Subset Drugbank file by 
 #DrugBank <-subset_GMT(DrugBank, DBoverlap_names)
  
# Subset to only drugs that have kinase targets
DrugBank_kinome <- GMT_kinome_filter(DrugBank, kinome)
  length(DrugBank_kinome)
  
# Write filtered GMT
dir = "Perturbation_Data/DrugBank/DrugBank_mechanisms.gmt"
if(file.exists(dir)){file.remove(dir) } else{print("Doesn't exist")}
gmtlist2file(CREEDS_manualDrugs_DRH, dir)  
```




# [Dataset 4] CREEDS_ManualGenes / kinome
```{r}
CREEDS_ManualGenes <- gmt2list("Perturbation_Data/CREEDS/single_gene_perturbations-v1.0.gmt")
CREEDS_ManualGenes_meta <- read.csv("Perturbation_Data/CREEDS/single_gene_perturbations-p1.0_metadata.csv")

# Filter to only kinase perturbations (same number of rows with either method)
## [Method 1: USING KINOME DATABASE]
## ****INCLUDE ROWS WITH KINASES IN RESPONSE AS WELL??
  creeds_genes <-substr(names(CREEDS_ManualGenes), 1 ,nchar(names(CREEDS_ManualGenes))-3)
  kinome_genes.filter <-toupper(creeds_genes) %in% toupper(kinome$Name)
  CREEDS_ManualGenes_kinome <-CREEDS_ManualGenes[kinome_genes.filter]

## [METHOD 2: USING CREEDS METADATA] CAN'T FILTER THIS WAY BC THEY DON'T SYSTEMATICALLY INCLUDE GENE SYMBOL IN THE METADATA....
  #kinome_genes.filter <-CREEDS_ManualGenes_meta[ grep("kinase",          
  #CREEDS_ManualGenes_meta$hs_gene_symbol), ] 
  #CREEDS_ManualGenes_kinome <-subset_GMT(CREEDS_ManualGenes, kinome_genes.filter$)

length(CREEDS_ManualGenes_kinome)
# Write filtered GMT
dir = "Perturbation_Data/CREEDS/Filtered_GMTs/CREEDS_ManualGenes_kinases.gmt"
if(file.exists(dir)){file.remove(dir) } else{print("Doesn't exist")}
gmtlist2file(CREEDS_ManualGenes_kinome, dir)  
```

## Kinase distribution metrics
```{r, fig.width=9, fig.height=9}
## Individual kinase frequecies
library(ggplot2)
# By Gene
perturbed_kinases <-substr(names(CREEDS_ManualGenes_kinome),
                           1 ,nchar(names(CREEDS_ManualGenes_kinome))-3)
kinase.table <- data.frame(Kinases = as.data.frame(table(perturbed_kinases))[,1],
                                Frequency = as.data.frame(table(perturbed_kinases))[,2]) 
ggplot(data=kinase.table, aes(x=Kinases,y=Frequency, fill = Kinases) ) + 
  geom_bar(stat="identity") + coord_polar(theta="x") + theme(legend.position="none") + labs(title="Kinase Frequencies: Gene-level")
```
```{r}
# By Group
Group <-kinome[kinome$Name %in% perturbed_kinases,]$Group
kinase.table <- data.frame(Group = as.data.frame(table(Group))[,1],
                                Frequency = as.data.frame(table(Group))[,2]) 
ggplot(data=kinase.table, aes(x=Group,y=Frequency, fill = Group) ) + 
  geom_bar(stat="identity", color="white") + coord_polar(theta="x") + labs(title="Kinase Frequencies: Group-level")
```
```{r}
# By Family
Family <-kinome[kinome$Name %in% perturbed_kinases,]$Family
kinase.table <- data.frame(Family = as.data.frame(table(Family))[,1],
                                Frequency = as.data.frame(table(Family))[,2]) 
ggplot(data=kinase.table, aes(x=Family,y=Frequency, fill = Family) ) + 
  geom_bar(stat="identity", color="white") + coord_polar(theta="x") + labs(title="Kinase Frequencies: Family-level")
```
## t-SNE plots
```{r}

```

# [Dataset 5] GEO_KinasePerturbations
```{r}
# Kinase perturbations from GEO (found on ENRICHR)
GEO_KinasePert.up <-gmt2list("Perturbation_Data/GEO/Kinase_Perturbations_from_GEO_up.gmt")
GEO_KinasePert.dn <-gmt2list("Perturbation_Data/GEO/Kinase_Perturbations_from_GEO_down.gmt")

length(GEO_KinasePert.dn); length(GEO_KinasePert.up)
```








# **[Dataset 6] LINCS_L1000_Chem**
## Filter out score next to genes in LINCS_chem data
```{r, val=F}
# Import LINCS L1000 chem perturbation dataset
#chem_up <-read.delim("Perturbation_Data/LINCS_L1000_Chem/LINCS_L1000_Chem_Pert_up.txt", header=F) # Look for kinase activator drugs
#chem_dn <-read.delim("Perturbation_Data/LINCS_L1000_Chem/LINCS_L1000_Chem_Pert_down.txt",header=F) # Look to kinase-inhibitor drugs

# Filter L1000 experiments with drugs whose mechanisms are kinase-related AND Match to expectations about how they operate

## Convert chem_up/dn to a recognizable format (get rid of "1.0")
#filter_LINCS_chem <- function(chem_pert){
#  geneLists = c()
#  lineNames = c()
#  genes_df = data.frame()
#  for(number in 1:nrow(chem_dn)){
#    line = chem_dn[number,]
#    line_name = c(t(line))[1]
#    gene_line = c(t(line))[-1]
#    gene_line = gene_line[as.character(gene_line)!=""]
#    geneLists = c(geneLists, paste(sub(",1.0","", gene_line), collapse="\t"))
#    lineNames =  c(lineNames, line_name)
#  }
#  return(list(lineNames, geneLists))
#}
#filter_chem_up <-filter_LINCS_chem(chem_up)
#genes_df = data.frame(lineNames=filter_chem_up[1], geneLists=filter_chem_up[2])
## Write to file
#write.table(test, "Perturbation_Data/LINCS_L1000_Chem/Chem_dn.edit.txt", sep #='\t', col.names = F, row.names = F, quote = F)
```

## Import edited chem files
```{r}
setwd("~/Desktop/X2K_Genetic_Algorithm/Validation/")
# Import kinome list
library(readxl)
kinome <- read_excel("../General_Resources/Kinase.com/Kinome_Hsap.xls")
#kinome_updated <- read_excel("~/Desktop/x2k/General_Resources/Kinase.com/Kinome_Hsap_updated.xls")
kea_kinases <- read.delim("../data/KEA/kea_ranks.txt", header=F, col.names = c("Name", "Rank1","Rank2"))

# IMPORT PYTHON-EDITED L1000 chem pert data
chem_up_edit <-read.delim("Perturbation_Data/LINCS_L1000_Chem/LINCS_L1000_Chem_Pert_up_editPY.txt", header=F)

chem_dn_edit <-read.delim("Perturbation_Data/LINCS_L1000_Chem/LINCS_L1000_Chem_Pert_dn_editPY.txt", header=F, colClasses = "character")
```


## DrugRepurposingHub filter
### All experiments separate
```{r}  
# **** HOW DO YOU TELL IF A DRUG IS INHIBITING OR ACTIVATING THE TARGET
# Import metadata from L1000FWD website
chem_meta <-read.csv("Perturbation_Data/LINCS_L1000_Chem/CD_signature_metadata.csv", header=T)[,c("sig_id","pert_desc")]
##  Kathleen's DrugCentral ID list is the most comprehensive (1500+ drugs) BUT it doesn't include HOW it affects it's targets (inhibit, activate) aka Mechanism of Action (MOA)
  #DrugCentral = read.csv("Drug-Target_Data/Drug Central/DrugCentral_EdgeList_v8_RoughManualCorrections_12-06-17.csv", header=T)
## Use Drug Repurposing Hub instead
DRH <-read.delim("Drug-Target_Data/Drug Repurposing Hub/Repurposing_Hub_export.txt", header=T) # All entries have TargeNames

#DRH_filter <-subset(DRH, DRH$Target %in% kinome$Entrez_Symbol) # 222
#DRH_filter <-subset(DRH, DRH$Target %in% toupper(kinome$Name))# 148, 167
#DRH_filter <-subset(DRH, DRH$Target %in% toupper(kinome_updated$Name))# 161
# DRH_filter <-subset(DRH, DRH$Target %in% kea_kinases$Name ) #253

# Filter DRH to only drugs that target kinases
  #DRH_filter = data.frame()
  #for(row_num in 1:dim(DRH)[1]){
  #  genesRow = DRH$Target[row_num]
  #  genes = strsplit(as.character(genesRow),", ")[[1]]
  #  if(sum(genes %in% kea_kinases$Name)>0 & genesRow!=""){
  #    DRH_filter = rbind(DRH_filter, DRH[row_num,])
  #  }
  #}
## Filter DRH to only include kinase inhibitors
### %like% returns all rows with the substring "kinase inhibitor" somewhere in the MOA col
library(dplyr); library(data.table)
DRH_kinaseInhibitors = subset(DRH, MOA %like% "kinase inhibitor") # 217
DRH_kinase.Antagonist = subset(DRH, MOA %like% "kinase antagonist") # = 0
DRH_kinase.Agonist = subset(DRH, MOA %like% "kinase agonist") # = 0

# Filter any instances with no gene targets listed
DRH_kinaseInhibitors = subset(DRH_kinaseInhibitors, Target!="") #201


### Get drug name from first col and make into new col
get_kinaseInhibitor_experiments <- function(chem_pert){
#chem_pert <- chem_up_edit
  Names=c(); DEGlist=c(); LINCS_DrugList=c()
  for(number in 1:nrow(chem_pert)){
    rowName = as.character(chem_pert[number,]$V1)
    rowSplit = strsplit(rowName,"-")[[1]]
    Drug = rowSplit[2:(length(rowSplit)-1)] # get first and last elements only
    Drug = paste(Drug,collapse="-") # Collapse if multiple elements still
    # Add to vector if drug is a kinase inhibitor
    if(toupper(Drug) %in% toupper(DRH_kinaseInhibitors$Name)){
      LINCS_DrugList = c(LINCS_DrugList, Drug)
      # Get kinase targets and put into the name of each experiment
      kTargets = strsplit(as.character(DRH_kinaseInhibitors[toupper(DRH_kinaseInhibitors$Name)==toupper(Drug),"Target"][1]),", ")[[1]]
      # Strip any invisible \n
      kTargets <-gsub("[\r\n]", "", kTargets)
      kinaseTargets <- paste("_[",paste(kTargets,collapse="|"),"]", sep="")
      Names = c(Names, paste(rowName,kinaseTargets,sep=""))
      # Get DEGs for each drug
      DEGcols = c(t(chem_pert[number,-1])) # Get all cols except first
      DEGs = paste(DEGcols[DEGcols!=""], collapse="\t")
      DEGlist = c(DEGlist, DEGs)
    }
  }
  return(list(Drug=LINCS_DrugList, Names=Names, DEGlist=DEGlist)) }
  
# CHEM_UP
up_output <-get_kinaseInhibitor_experiments(chem_up_edit)
chem_up_kinaseInhibitors <- data.frame(Drug=up_output$Drug, 
                                       Name=up_output$Names, ExtraTab="",
                                       Genes=up_output$DEGlist)
dim(chem_up_kinaseInhibitors)
## Write file
write.table(chem_up_kinaseInhibitors[,-1], "Perturbation_Data/LINCS_L1000_Chem/DrugRepurposingHub_filtered/Chem_up_DRH.kinaseInihibitors.txt", sep='\t',col.names = F, row.names = F, quote=F)

# CHEM_DN
dn_output = get_kinaseInhibitor_experiments(chem_dn_edit)
chem_dn_kinaseInhibitors <- data.frame(Drug=dn_output$Drug,
                                      Name=dn_output$Names, ExtraTab="",
                                      Genes=dn_output$DEGlist)
dim(chem_dn_kinaseInhibitors)
## Write file
write.table(chem_dn_kinaseInhibitors[,-1], "Perturbation_Data/LINCS_L1000_Chem/DrugRepurposingHub_filtered/Chem_dn_DRH.kinaseInihibitors.txt", sep='\t',col.names = F, row.names = F, quote=F)
```

### Collapse same drugs into single rows
```{r}
length(unique(chem_up_kinaseInhibitors$Drug)) # Number of unique kinase inhibitor drugs (n=46)

# chem_pert_kinaseInhibitors=chem_dn_kinaseInhibitors
collapse_by_drugs <- function(chem_pert_kinaseInhibitors){
  library(dplyr)
  collapsed_geneList=c(); collapsed_NameList=c()
  for(drug in unique(chem_pert_kinaseInhibitors$Drug)){
    drug_subset <-chem_up_kinaseInhibitors[chem_up_kinaseInhibitors$Drug==drug,]
    
    # Collapse DEGs (get average ranks to avoid bias with repeating genes)
    geneName=c(); geneRank=c()
    for(list in drug_subset$Genes){
      geneSplit <- strsplit(list,"\t")[[1]]
      geneName <- c(geneName, geneSplit)
      geneRank <- c(geneRank, 1:length(geneSplit))
    }
    ranking_df <-data.frame(Gene.Name=geneName, Gene.Rank=geneRank)
    # Get average ranking
    average_rank <- ranking_df %>% group_by(Gene.Name) %>% dplyr::summarise(meanRank=mean(Gene.Rank)) %>% as.data.frame()
    # Sort genes by rank
    sorted_genes <- average_rank[order(average_rank$meanRank), ]$Gene.Name %>% as.character() %>% paste(collapse="\t")
    # Create new row name
    # Get name elements
    LINCS_ID <- strsplit(as.character(drug_subset$Name[1]),"_")[[1]][1]
    targets <- strsplit(as.character(drug_subset$Name)[1],"_")[[1]][4]
    newName = paste(LINCS_ID, drug, targets, sep="_")
    collapsed_NameList <- c(collapsed_NameList, newName)
    collapsed_geneList <- c(collapsed_geneList, sorted_genes)
  }
  chem_up_kinaseInhibitors.collapse <- data.frame(Name=collapsed_NameList,
                                                  ExtraTab="",
                                                  Genes=collapsed_geneList)
  return(chem_up_kinaseInhibitors.collapse)
}

# CHEM_UP COLLAPSE
chem_up_kinaseInhibitors.collapse = collapse_by_drugs(chem_up_kinaseInhibitors)
## Write to file
write.table(chem_up_kinaseInhibitors.collapse, "Perturbation_Data/LINCS_L1000_Chem/DrugRepurposingHub_filtered/Chem_up_DRH.kinaseInihibitors_collapsed.txt", sep='\t',col.names = F, row.names = F, quote=F)

# CHEM_DN COLLAPSE
chem_dn_kinaseInhibitors.collapse = collapse_by_drugs(chem_dn_kinaseInhibitors)
## Write to file
write.table(chem_dn_kinaseInhibitors.collapse, "Perturbation_Data/LINCS_L1000_Chem/DrugRepurposingHub_filtered/Chem_dn_DRH.kinaseInihibitors_collapsed.txt", sep='\t',col.names = F, row.names = F, quote=F)

```

## KinomeScan filter
### Import KinomeScan datasets
```{r}
# curl -H "Accept: text/csv" http://lincs.hms.harvard.edu/db/api/v1/protein/ >> prot.json
# curl -H "Accept: text/csv limit: inf" http://lincs.hms.harvard.edu/db/api/v1/protein/ >> prot.json
#as.data.frame(out)
library(jsonlite); library(readxl); #library(svMisc)
expts <-as.data.frame(read_excel("Drug-Target_Data/Kinome_Scan/LINCS_HMS_ExperimentList.xlsx"))
datasetIDs <- subset(expts, `HMS Dataset Type`=="KINOMEscan" )$`HMS Dataset ID` #|`HMS Dataset Type`=="KiNativ"
length(datasetIDs)

# Import LINCS Drug IDs
# Get small molecule (sm) info
#scID <- unique(KS_expt_sub$smallmolecule_smCenterCompoundID)
#saltID <- unique(KS_expt_sub$smallmolecule_smSalt)
  #url_sm <- paste("http://lincs.hms.harvard.edu/db/api/v1/","smallmolecule/",scID,"-", saltID, "/", sep="")
  #KS_sm <-jsonlite::fromJSON(url_sm)
```

### Create GMT with drug names and DEGS ordered by rank
```{r}
## SETUP FUNCTION FOR NEXT PART
create_Drug.KinaseTarget_df <- function(KS_expt_sub, decreasing){
 # Convert to numeric
    KS_expt_sub$datapointValue <- as.numeric(KS_expt_sub$datapointValue)
    # Order df according to binding 
    KS_expt_sub.sort <- KS_expt_sub[order(KS_expt_sub$datapointValue, decreasing=decreasing),]
    # Rescale/normalize binding affinity
    #KS_expt_sub.sort$datapointValue<-rescale(KS_expt_sub.sort$datapointValue, c(0,1))
    ## Get drug name
    DrugSalt = paste(unique(KS_expt_sub.sort$smallmolecule_smLincsID),
                     unique(KS_expt_sub.sort$smallmolecule_smName),
                     unique(KS_expt_sub.sort$smallmolecule_smSalt),ID,sep="_")
    hist(KS_expt_sub.sort$datapointValue)
    ## Get kinases with respective scores
    kinase_list = c()
    for(row_num in 1:nrow(KS_expt_sub.sort)){
      kinase = KS_expt_sub.sort[row_num,]$protein_ppName
      
      #kinase_score = KS_expt_sub.sort[row_num,]$datapointValue
      # Get GeneSymbols for each kinase
      if(kinase %in% prots$Name){
      GeneSymbol <- prots[prots$Name==kinase,]$Gene.Symbol[1]
      }
      else{GeneSymbol<-kinase}
      # Compile into list of kinases with scores
      kinase_list <- paste(kinase_list, paste(GeneSymbol, sep=","), "\t", sep="")
    }
    new_row = data.frame(Name=DrugSalt, ExtraTab="",Kinases=kinase_list)
    #Drug_KinaseTargets <- rbind(Drug_KinaseTargets, data.frame(Name=DrugSalt, Kinases=kinase_list))
    return(new_row)
    }
  


# GET DATA FROM APIs and organize into df
#get_Drug_KinaseTargets <- function(datasetIDs){
  library(scales)
  Drug_KinaseTargets = data.frame()
  unrecovered_IDs = c()
  # Import HMS LINCS protein metadata
  prots <-data.frame(read_excel("Drug-Target_Data/Kinome_Scan/proteins_20180116225538.xlsx"))
  prots <- prots[!is.na(prots$Gene.Symbol),]
  pbar=1
  for(ID in datasetIDs){
    #pbar =  pbar+1
    #progress(pbar, progress.bar=T) # Animate progress bar
    datasetID <-ID #"20211"
    url <-paste("http://lincs.hms.harvard.edu/db/api/v1/","datasetdata/", datasetID, sep="")
    KS_expt <-jsonlite::fromJSON(url)
    
    if(length(KS_expt) != 0){
      KS_expt <- KS_expt[KS_expt$datapointValue!="ND",] # Get rid if NA values
      KS_expt$datapointValue <- gsub(">", "", KS_expt$datapointValue) 
    # Get just binding affinity results
    # "dissociationConstant": Higher value means the more strongly the drug binds to the kinase'
    # "percentControl": Higher value means the more strongly the drug binds to the kinase?
      if("percentControl" %in% KS_expt$datapointName){
        print(paste("*percentControl* dataset",ID))
        KS_expt_sub <-subset(KS_expt, datapointName=="percentControl")
        Drug_KinaseTargets <- rbind(Drug_KinaseTargets, 
                                    create_Drug.KinaseTarget_df(KS_expt_sub,
                                                                decreasing=T))
        }
      else if("dissociationConstant" %in% KS_expt$datapointName){
        print(paste("*dissociationConstant* dataset",ID))
         KS_expt_sub <-subset(KS_expt, datapointName=="dissociationConstant")
          Drug_KinaseTargets <- rbind(Drug_KinaseTargets, 
                                    create_Drug.KinaseTarget_df(KS_expt_sub,
                                                                decreasing=T))
          }
      else if("percentInhibition" %in% KS_expt$datapointName){
        print(paste("*percentInhibition* dataset:",ID))
        KS_expt_sub <-subset(KS_expt, datapointName=="percentInhibition")
        Drug_KinaseTargets <- rbind(Drug_KinaseTargets, 
                                    create_Drug.KinaseTarget_df(KS_expt_sub,
                                                                decreasing=T))
        }
      else{print(paste("Can't find appropriate measure for ID:",ID)); 
      unrecovered_IDs = c(unrecovered_IDs, ID) }
    }
    else{ print(paste("No data recovered for ID:",ID)); 
    unrecovered_IDs = c(unrecovered_IDs, ID) }
    }  
  
  #return(Drug_KinaseTargets) }
#Drug_KinaseTargets <-get_Drug_KinaseTargets(datasetIDs)
write.table(Drug_KinaseTargets, "Drug-Target_Data/Kinome_Scan/KinomeScan_Drug.KinaseTargets_noScores.txt", sep='\t',col.names = F, row.names = F, quote=F)
```

#### Collapse KINOMEscan data by drug, Get average rank for each DEG
```{r}
# Get list of drugs in KINOMEscan and put back in df
KINOMEscan_drugs = c()
for(rowname in Drug_KinaseTargets$Name){
  KINOMEscan_drugs = c(KINOMEscan_drugs, strsplit(as.character(rowname),"_")[[1]][2])
}
Drug_KinaseTargets$Drug <- KINOMEscan_drugs


# COLLAPSE DF BY NAME
library(dplyr)
averaged.GeneLists=c(); averaged.NameList=c(); KS.filtered_DrugList=c()
for(drug in unique(Drug_KinaseTargets$Drug)){
    KS.filtered_DrugList <- c(KS.filtered_DrugList, drug)
    drug_subset <- subset(Drug_KinaseTargets, Drug==drug)
      newName <- paste(strsplit(as.character(drug_subset$Name[1]),"_")[[1]][1:2],collapse="_")
  
      # Get average of all genes
      for(list in drug_subset$Kinases){
          averaged_rank.subset=c()
          geneSplit <- strsplit(list,"\t")[[1]]
          geneRank <-1:length(geneSplit)
          ranking_df <-data.frame(Gene.Name=geneSplit, Gene.Rank=geneRank)
          # Get average ranking
          average_rank <- ranking_df %>% group_by(Gene.Name) %>% dplyr::summarise(meanRank=mean(Gene.Rank)) %>% as.data.frame()
        # Sort genes by rank
        sorted_genes <- average_rank[order(average_rank$meanRank), ]$Gene.Name %>% as.character() %>% paste(collapse="\t")
        # Compile genes into subset
        averaged_rank.subset <- paste(averaged_rank.subset, sorted_genes, collapse='\t')
      }
  averaged.GeneLists <- c(averaged.GeneLists, averaged_rank.subset)
  averaged.NameList <- c(averaged.NameList, newName)
  }

Drug_KinaseTargets.avgRank <-data.frame(Name=averaged.NameList, ExtraTab="",
                                        rankedDEGs=averaged.GeneLists,
                                        Drug=KS.filtered_DrugList)
dim(Drug_KinaseTargets.avgRank)
    

# Write file
write.table(Drug_KinaseTargets.avgRank[-4], "Drug-Target_Data/Kinome Scan/Drug_KinaseTargets.avgRank.txt", sep='\t',col.names = F, row.names = F, quote=F)
```

### Filter LINCS L1000 by drugs that appears in KINOMEscan
```{r}
# Drugs in LINCS L1000
  # chem_pert = chem_up_edit
subset_L1000_by_KINOMEscan <-function(chem_pert){
  newName_list=c(); filteredDEG_list=c()
  for(row.num in 1:nrow(chem_pert)){
      rowName = as.character(chem_pert[row.num,]$V1)
      rowSplit = strsplit(rowName,"-")[[1]]
      L1000_Drug = rowSplit[2:(length(rowSplit)-1)] # get first and last elements only
      L1000_Drug = paste(L1000_Drug,collapse="-") # Collapse if multiple elements still
      if(toupper(L1000_Drug) %in% toupper(Drug_KinaseTargets.avgRank$Drug)){
        # Use L1000 DEGs
        DEGs = c(t(chem_pert[row.num,-1]))
        DEGs = DEGs[DEGs!=""]
        filteredDEG_list = c(filteredDEG_list, paste(DEGs,collapse="\t"))
        # Create new Name from L1000 name + unbiased "targets" from KINOMEscan
        kinomescan_targets <- Drug_KinaseTargets.avgRank[toupper(Drug_KinaseTargets.avgRank$Drug)==toupper(L1000_Drug),]$rankedDEGs
        kinomescan_targets.mod <- paste(strsplit(as.character(kinomescan_targets),"\t")[[1]],collapse="|")
        newName <- paste(rowName,paste("[",trimws(kinomescan_targets.mod),"]",sep=""),sep="_")
        newName_list <- c(newName_list, newName)
      }
  }
  # Create new 
  return(list(newName_list=newName_list, filteredDEG_list=filteredDEG_list))
}

# CHEM_UP 
L1000_KINOMEscan.filter.UP <-subset_L1000_by_KINOMEscan(chem_up_edit)
L1000_KINOMEscan.filter.UP.df <- data.frame(Names=L1000_KINOMEscan.filter.UP$newName_list, ExtraTab="",
                                  DEGs=L1000_KINOMEscan.filter.UP$filteredDEG_list)
write.table(L1000_KINOMEscan.filter.UP.df,"Perturbation_Data/LINCS_L1000_Chem/KinomeScan_filtered/LINCS-L1000_KINOMEscan.UP_filtered.txt", sep='\t',col.names = F, row.names = F, quote=F)
# CHEM_DN
L1000_KINOMEscan.filter.DN <-subset_L1000_by_KINOMEscan(chem_dn_edit)
L1000_KINOMEscan.filter.DN.df <- data.frame(Names=L1000_KINOMEscan.filter.DN$newName_list, ExtraTab="",
                                  DEGs=L1000_KINOMEscan.filter.DN$filteredDEG_list)
write.table(L1000_KINOMEscan.filter.DN.df,"Perturbation_Data/LINCS_L1000_Chem/KinomeScan_filtered/LINCS-L1000_KINOMEscan.DN_filtered.txt", sep='\t',col.names = F, row.names = F, quote=F)
```




# Summaries
```{r, echo=F}
print("*****************SINGLE DRUG PERTURBATIONS*****************")

# CREEDS_single_drug_perturbations / DrugBank.TargetCentral + kinome
print(paste("# of CREEDS_manualDrugs rows =", length(CREEDS.ManualDrugs) ))
print(paste("# of CREEDS_manualDrugs rows, that have kinase targets, and are in DrugBank.TargetCentral list = ",length(CREEDS_manualDrugs_DB.TC)))
# CREEDS_single_drug_perturbations / DrugRepurposingHub + kinome
print(paste("# of CREEDS_manualDrugs rows =", length(CREEDS.ManualDrugs) ))
print(paste("# of CREEDS_manualDrugs rows, that have kinase targets and are in DrugRepurposingHub list = ",length(CREEDS_manualDrugs_DRH)))

# CREEDS_DrugMatrix / DrugBank.TargetCentral + kinome
print(paste("# of CREEDS_DrugMatrix rows =", length(CREEDS_DrugMatrix) ))
print(paste("# of CREEDS_manualDrugs rows, that have kinase targets,03 and are in DrugBank.TargetCentral list = ",length(CREEDS_DrugMatrix_DB.TC)))
# CREEDS_DrugMatrix / DrugRepurposingHub + kinome
print(paste("# of CREEDS_DrugMatrix rows =", length(CREEDS_DrugMatrix) ))
print(paste("# of CREEDS_DrugMatrix rows, that have kinase targets, and are in DrugRepurposingHub list = ",length(CREEDS_DrugMatrix_DRH)))

# DrugBank_mechanisms / kinome
print(paste("# of DrugBank (Moshe) rows =", length(DrugBank) ))
print(paste("# of DrugBank (Moshe) rows, that have kinase mechanisms in DrugBank = ",length(DrugBank_kinome)))



print("*****************SINGLE GENE PERTURBATIONS*****************")
# CREEDS_ManualGenes filtered by kinases
print(paste("# of CREEDS_ManualGenes rows =", length(CREEDS_ManualGenes)))
print(paste("# of CREEDS_ManualGenes rows, that ARE kinases =", length(CREEDS_ManualGenes_kinome)))
#GEO_KinasePerturbations
print(paste("# of GEO_KinasePerturbation rows (up/down) =", length(GEO_KinasePert.dn)*2))


```


